---
title: "Plotting the contours plot to study the joint distribution of paired samples"
output: rmarkdown::html_vignette
author:
- "Kevin Ortiz Gonzalez"
- "Karol Michelle Sandoval"
date: "April 2022"
vignette: >
  %\VignetteIndexEntry{Plotting the contours plot to study the joint distribution of paired samples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The contour plot allows visualizing the joint distribution $f(x,y)$, in addition to being able to identify the possible existence of correlation between the distributions under study, which can be identified according to the slope of the contour plot. 

## Example 1

After running the BNP.test function to identify the existence of differences between the marginal distributions and subsequently the shift function to determine in which range such differences existed, we proceed to run the contour plot for the simulated data.

```{r contours1}
library(BNPPairedSamples) # Package library load

contours.plot(results_BNP = test_results)
```

In this case it is observed that the joint distribution is made up of two clusters and that in the vicinity of 0 the density is 0, also noteworthy is the increasing slope of the contours, which would be expressing a positive correlation between the distributions.

## Example 2

For the air quality data the differences were found visually in the location parameter, but when the shift was obtained it was observed that the 0 was contained in the interval and the differences observed visually were not significant.


```{r contours2}

contours.plot(results_BNP = test_results)
```

The contours show a flat shape, which would indicate that there is no correlation between marginal distributions 1 and 2. 
